---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setwd("D:/Coursera/Reproducible Research/Project 1/")
library('plyr')
library('ggplot2')
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
```


## What is mean total number of steps taken per day?
```{r}
sumDf <- ddply(df,.(date),summarize, sum=sum(steps))
plot(sumDf, xlab='Date',ylab='Steps Taken',type='h')
```


## Mean and the median total number of steps taken per day

```{r}
mean(df$steps,na.rm = TRUE)
median(df$steps,na.rm = TRUE)
```


## What is the average daily activity pattern?


```{r}
timeSeries <- ddply(df,.(interval),summarize,mean=mean(steps,na.rm = TRUE))
plot(timeSeries,type='l',ylab='Steps Taken')
```


## Interval with highest number of steps


```{r}
timeSeries[which.max(timeSeries$mean),1]
```


## Imputing missing values

1. Number of Missing values 

```{r}
sum(is.na(df$steps))
```


2. Filling in all the missing values


```{r}
meanDf <- ddply(df,.(interval),summarize,mean=mean(steps,na.rm = TRUE))
df2 <- merge(df,meanDf,by="interval",sort = TRUE)
```


3. Create a new Dataset with missing value filled in


```{r}
df2[is.na(df2$steps),2] <- df2[is.na(df2$steps),4]
df2 <- df2[with(df2, order(date)),]
df2 <- df2[,c("steps","date","interval")]
```


4. New Histogram and Mean & Median


```{r}
histDf <- ddply(df2,.(date),summarize, sum=sum(steps))
plot(histDf,xlab="Date", ylab="Steps Taken",type="h")
mean(df2$steps)
median(df2$steps)
```


## Are there differences in activity patterns between weekdays and weekends?
1. Creating a factor variable "daytype"

```{r}
df2$daytype <- ifelse((weekdays(df2$date, abbreviate = FALSE)) %in% c("Saturday","Sunday"), "weekend","weekday")
```


2. Make a panel plot containing a time series plot 


```{r}
aggDf <- ddply(df2,.(daytype,interval),summarize,avg = mean(steps))
qplot(interval,avg,data = aggDf, facets= daytype~.,geom = "line", xlab = "Interval", ylab = "Steps Taken")
```